### 2020-03-23 3_implementation.tex
Rewrite implementation to save space.

We have implemented Task Bench in the 15 parallel and distributed
programming systems listed in Table~\ref{tab:systems}. These include
traditional HPC programming models: MPI~\cite{MPI} and MPI+X (OpenMP
and CUDA); PGAS and actor models: Chapel~\cite{Chapel15},
Charm++~\cite{Charmpp93}, X10~\cite{X1005}; and task-based
systems: OmpSs~\cite{OmpSs11}, OpenMP 4.0~\cite{OpenMPSpec40}, PaRSEC~\cite{PARSEC13, PARSEC_DTD}, Realm~\cite{Realm14},
Regent~\cite{Regent15}, and StarPU~\cite{StarPU11}. Also, we consider
several systems used in large scale data analytics and workflows:
Dask~\cite{Dask15}, Spark~\cite{Spark10}, Swift/T~\cite{Wozniak13},
and TensorFlow~\cite{TensorFlow15}. We describe the systems and implementations below.

@@

\subsection{Chapel}

Chapel~\cite{Chapel15} is a parallel programming language
with a \emph{multi-res\-o\-lu\-tion} approach, supporting parallelism at a variety of
levels. Chapel's core features are a partitioned global address space
(PGAS), data distributions, tasks,
synchronization primitives, and array promotion. For Task Bench, we target a low level of
Chapel, using explicit task instantiation (via
\lstinline[language=Chapel]{coforall}), bulk access to distributed
arrays for efficient data movement, and atomic integers for
synchronization.

\subsection{Charm++}

Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors
(\emph{chares}) do not share memory and communicate via messages. Our Task Bench implementation uses a chare
array for the task graph, with one chare for each column, and uses messages to implement dependencies.

%% Collectives synchronize the
%% completion of the task graph execution.

% FIXME: Discuss SMP vs non-SMP versions

\subsection{Dask}

Dask~\cite{Dask15} is an implicitly parallel task-based programming system for large
scale data analytics with Python. Dask provides abstractions over
distributed NumPy arrays and Pandas dataframes, and also a lower-level
interface for launching tasks. We implement Task Bench in the lower-level interface.

\subsection{MPI and MPI+X}

We provide two MPI~\cite{MPI} implementations, written in the style of
communicating sequential processes (CSP) and bulk synchronous
programming (BSP). An excerpt from the CSP
implementation is shown in Listing~\ref{lst:code-sample}. The BSP
implementation uses \lstinline[language=C++]{MPI_Barrier} to enforce
boundaries between phases.

@@

\subsection{OpenMP and OmpSs}

OpenMP~\cite{OpenMPSpec40} and OmpSs~\cite{OmpSs11} are programming
models for loop- and task-based parallelism. (OmpSs is
source-compatible with OpenMP.) Our implementations use OpenMP 4.0
tasking.  Because tasks have a fixed number of dependencies,
we use a switch statement to implement dynamic dependencies for Task
Bench. For OpenMP, we tested GNU GOMP and Intel KMP and found KMP to be better
performing.

% FIXME: More to say about OmpSs? Is there anything different vs the OpenMP implementation?

%\subsection{OpenMP}

%OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
%parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
%for OpenMP is similar to OmpSs and uses
%task dependencies. We tested with GNU GOMP and Intel KMP
%and found KMP to be better performing.

\subsection{PaRSEC}

PaRSEC supports two distinct
programming models: \emph{parameterized task graphs}
(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
(DTD)~\cite{PARSEC_DTD}.  PTG is a dataflow model in which users
write a concise, algebraic description of the tasks and dataflows in
the program. This compressed representation is expanded into a full
task graph by a source-to-source compiler.

In DTD, tasks are enumerated in program order and dependencies are
identified automatically based on task arguments. To improve
scalability, the program is executed in SPMD fashion on all nodes, and
the user is responsible for eliding tasks that are not direct
predecessors or successors to those executed on the current node.

We implement Task Bench with both PTG and DTD, as well as a third
option based on DTD that has additional manual optimizations applied
to further reduce the cost of dynamic checks for task elision.

%% Task Bench uses a 2D block cyclic data distribution provided by PaRSEC. Tiles of data are 
%% distributed across nodes in pre-defined patterns,
%% and can be passed to tasks on other nodes, but
%% can only be written on the node where the data is originally
%% allocated. In our Task Bench implementation, each task writes to a
%% tile as output, and these tiles are passed as inputs to the
%% subsequent sets of tasks. As in the OmpSs and OpenMP implementations,
%% the task launch code uses a switch statement as the PaRSEC API assumes
%% tasks have a fixed number of dependencies.

\subsection{Realm}

Realm~\cite{Realm14} is an explicitly-parallel task-based programming
model used internally by Regent~\cite{Regent15}. Realm is thus a limit
study of what can be achieved with Regent.

Tasks in Realm are asynchronous, with dependencies
defined by \emph{events} passed from one task
to another. Realm's data model supports collections that live in a
specific memory. Data must be explicitly copied, and the copies
synchronized with tasks via events.
Realm also supports a \emph{subgraph} API that enables additional
optimizations. We use subgraphs in the Task
Bench implementation.

\subsection{Regent}

Regent~\cite{Regent15} is an implicitly-parallel task-based language
which implements the Legion programming model~\cite{Legion12}. We use
Regent rather than Legion since the Regent compiler
provides a critical optimization to
improve scalability~\cite{ControlReplication17}.

\subsection{Spark}

Spark~\cite{Spark10} is an implicitly-parallel programming model for
data analytics.
The core abstractions in Spark are functional operators such as map,
reduce, and join. Data in Spark are stored in
\emph{resilient
  distributed datasets} (RDDs): globally-visible,
dynamically single-assignment data structures. Spark caches RDDs in memory to avoid unnecessary
disk traffic.

Spark has tasks internally but they are not exposed to the user,
so a Task Bench implementation must map the task graph to operators that result in the
desired execution. We use
\lstinline[language=Scala]{flatMap} and
\lstinline[language=Scala]{groupByKey} to generate
dependencies, and then \lstinline[language=Scala]{mapPartitions} to execute a
series of tasks. An explicit hash partitioner ensures that
Spark does not attempt to group multiple Task Bench tasks into a
single Spark task, as tasks in Task Bench already represent coarse-grained units of work.

We performed extensive experiments to verify that no
extraneous factors interfered with our Spark measurements. We
disabled logging, ensured no RDDs are written
to disk, confirmed there is no measurable overhead due to JNI
calls from Java to C or due to serializing data, and cross-checked results with known cases
with optimal scheduler throughput.

\subsection{StarPU}

StarPU~\cite{StarPU11} is a task-based system utilizing a \emph{sequential task flow}
programming model similar to PaRSEC's DTD. Our Task Bench implementation in
StarPU is very similar to PaRSEC.
%% We use a simplified version of the 2D block cyclic data distribution from the Chameleon 
%% project~\cite{Chameleon}.

\subsection{Swift/T}

Swift/T~\cite{Wozniak13} is a parallel scripting language intended
primarily for the composition of HPC workflows. Swift/T
programs follow dataflow semantics, where every statement may
potentially execute in parallel as soon as its dependencies are
satisfied; statements only execute sequentially when explicitly
requested. The Swift/T compiler performs a number of optimizations to
improve performance of highly parallel programs~\cite{Armstrong14}. 
Our Task Bench in Swift/T uses its dataflow
semantics to capture dependencies on other tasks.

\subsection{TensorFlow}

TensorFlow~\cite{TensorFlow15} is designed for deep learning
workloads. Although TensorFlow's API exposes machine learning
concepts, internally TensorFlow is a task graph execution engine,
making it a good fit for Task Bench. TensorFlow programs operate via
explicit graph construction. Task graphs are composed in Python and
run by a C++ execution engine.

\subsection{X10}

X10~\cite{X1005} is an explicitly parallel programming language for
place-based programming. Features include \emph{places}
which represent distributed memories, a PGAS model where references to
remote objects can be held (but can only be dereferenced on the local
place), asynchronous tasks, and a place-changing construct to move the
execution of a task to a remote place. X10 also supports a variety of
synchronization primitives.

Our Task Bench implementation uses
\lstinline[language=X10]{Rail.asyncCopy} for efficient data movement
between places, along with place-changing and atomic integers for
synchronization. We use the native backend of X10, which compiles to
C++.

>>>>

We have implemented Task Bench in the 15 parallel and distributed
programming systems listed in Table~\ref{tab:systems}. These include
traditional HPC programming models (MPI and MPI+X), PGAS and actor
models (Chapel, Charm++ and X10), task-based systems (OmpSs, OpenMP
4.0, PaRSEC, Realm, Regent, and StarPU) and systems for large scale
data analytics, machine learning and workflows (Dask, Spark, Swift/T,
and TensorFlow). We briefly describe the systems and implementations
below.

@@

\subsection{Traditional HPC Programming Models}

MPI~\cite{MPI} is a message-passing API for HPC. We provide an MPI
implementation written in the style of communicating sequential
processes (CSP). An excerpt of this implementation is shown in
Listing~\ref{lst:code-sample}.

@@

\subsection{PGAS and Actor Models}

PGAS and actor models, such as Chapel~\cite{Chapel15},
Charm++~\cite{Charmpp93} and X10~\cite{X1005}, provide support for
asynchronous tasks that make them amenable to a straightforward
implementation of task bench. Synchronization is explicit and may be
provided by messages (in actor models) or other primitives, such as
locks or atomics (in PGAS models). PGAS models such as Chapel and X10
provide global references to data anywhere in the machine, but vary in
whether data can be accessed remotely or not (Chapel allows this, X10
does not).

Chapel also provides support for implicit parallelism which we do not
evaluate in this paper.

\subsection{Task-Based Programming Models}

Task-based systems include OmpSs~\cite{OmpSs11}, OpenMP
4.0~\cite{OpenMPSpec40}, PaRSEC~\cite{PARSEC13, PARSEC_DTD},
Realm~\cite{Realm14}, Regent~\cite{Regent15}, and
StarPU~\cite{StarPU11}. Though the details vary, these systems
typically provide implicit parallelism, where tasks are enumerated
sequentially (in program order) and the dependencies between tasks are
analyzed automatically to construct a \emph{dependence graph} that
guides the execution of tasks. PaRSEC provides two modes,
\emph{dynamic task discovery} (DTD)~\cite{PARSEC_DTD}, which operates
as above, and \emph{parameterized task graphs} (PTG)~\cite{PARSEC13},
where the dependence graph is constructed automatically from an
analytical representation of the task graph. Realm (unlike the others
above) is explicitly parallel and requires tasks to be connected
explicitly via \emph{events}.

Among the distributed task-based systems, one question is how to
resolve the inherent scalability bottleneck posed by enumerating tasks
sequentially. PaRSEC and StarPU resolve this by permitting the user to
omit tasks that are not directly connected in the dependence
graph. The program is replicated (for efficiency of the analysis), but
each task is executed only once. Because the dynamic checks to see if
a task should be executed are not free of cost, we also provide a
PaRSEC implementation (under the label \lstinline{shard}) which has
been hand-written to avoid all such checks. Regent performs an
equivalent optimization at compile time~\cite{ControlReplication17}
that does not require user intervention.

\subsection{Data Analytics, Machine Learning and Workflows}

Dask~\cite{Dask15}, Spark~\cite{Spark10} and
TensorFlow~\cite{TensorFlow15} are programming models for large scale
data analytics and machine learning. Dask and TensorFlow provide
domain-specific abstractions built on top of task-based runtimes. Our
implementations directly create tasks and are similar to other
task-based systems above.

Spark provides support for functional operators that implicitly map to
tasks. We use \lstinline[language=Scala]{flatMap} and
\lstinline[language=Scala]{groupByKey} to generate dependencies, and
then \lstinline[language=Scala]{mapPartitions} to execute a series of
tasks. An explicit hash partitioner ensures the correct task
granularity.

Swift/T~\cite{Wozniak13, Armstrong14} is a parallel scripting language
with dataflow semantics, used primarily for workflow automation. Our
implementation is straightforward.
